Relative Stable Maps in Genus One via Central Alignments

### 2018-05-10 genus-one-relative-maps.tex
finished chen-kim review

We recall a number of definitions from \cite{ChenDegeneration}; see also \cite{KimLog}.

@@

Finally, fix an $n$-tuple of non-negative contact orders $(c_1,\ldots,c_n)$.
\begin{definition}
 A \emph{relative log prestable map} with tangency condition $\mathbf{c}$ is given by a diagram of log morphisms:
 \bcd
 ((C,\mathbf p),\pM_C)\ar[rr,"f"]\ar[dr] & & ((W,D),\pM_W)\ar[rr,"\pi_X"]\ar[dl] & & (X,Y)\times S \\
  & (S,\pM_S) & & &
 \ecd
 \begin{enumerate}
  \item $(C,\mathbf p),\pM_C)\to (S,\pM_S)$ is a minimal log twisted curve;
  \item $((W,\pM_W)\to (S,\pM_S),(W,D)\to(X,Y))$ is an extended twisted log expansion of $(X,Y)$;
  \item (\emph{corank condition}) \sidenote{split the definition into a more general one over LogSch and minimality?} for every geometric point $\bar{s}\in S$, the rank of \[\operatorname{Coker}(\pM^{W/S}_{S,\bar{s}}\to \pM_{S,\bar{s}})\] is equal to the number of non-distinguished nodes of $C_{\bar{s}}$; recall that a node of $C_{\bar{s}}$ is \emph{distinguished} if it is mapped under $f$ to the singular locus of $W_{\bar{s}}$.
  \item (\emph{log admissibility}) the morphism of log structures $f^\flat\colon f^*\pM_W\to\pM_C$ is simple at every distinguished node;
  \item (\emph{tangency condition}) the underlying map $\underline{f}\colon C\to (W,D)$ is non-degenerate, namely no component of $C$ is mapped entirely into $D$, and locally around every marking $p_i$ we can find a chart
  \bcd
  \NN\ar[r,"\cdot c_i"]\ar[d] & \NN\ar[d] \\
  f^*\pM^D\ar[r] & \pM^{p_i}
  \ecd
  for the restriction of $f^\flat$ to the $f^*\pM^D$ factor.
  \end{enumerate}
\end{definition}
\begin{definition}
 A relative log prestable map as above is \emph{stable} if for every geometric point $\bar{s}\in S$ the group of automorphisms $(\sigma,\tau)$ is finite, where:
 \begin{enumerate}
  \item $\sigma$ is an automorphism of $((C,\pM_C)\to (S,\pM_S))_{\bar{s}}$ preserving the markings $\mathbf p$;
  \item $\tau$ is an automorphism of $((W,\pM_W)\to (S,\pM_S))_{\bar{s}}$ preserving $D$ and $\pi_X\colon W_{\bar{s}}\to X$;
  \item $\tau\circ f_{\bar{s}}=f_{\bar{s}}\circ \sigma$.
 \end{enumerate}
\begin{remark}
 The corank and simplicity condition imply the following: the rank of the locally free log structure $\pM_S$ at $\bar{s}$ amounts to the number of irreducible components of $W_{\bar{s}}^{\text{sing}}$ plus the number of non-distinguished nodes of $C_{\bar{s}}$; all the smoothing parameters of the distinguished nodes mapping to the same component of $W_{\bar{s}}^{\text{sing}}$ have been identified \emph{up to twisting}. In facts, at a distinguished node $q$ of $C$, letting $R$ be the local ring of $S$ at $\pi(q)$, we can find \'{e}tale local charts
 \bcd
 \NN^2\oplus_{\NN} \pM_{S}\ar[r,"{(id,\cdot m_q)}"]\ar[d] & \NN^2\oplus_{\NN} \pM_{S}\ar[d] \\
 f^*\pM_W \ar[r,"f^\flat"]\ar[d] & \pM_C\ar[d] \\
 R[z_1,\ldots,z_{\dim(X)-1},x,y]/xy-t\ar[r,"\left(\substack{x\mapsto a^{m_q}\\ y\mapsto b^{m_q}\\ t\mapsto s^{m_q}}\right)" below=.2cm] & R[a,b]/ab-s
 \ecd
 where the twist at $f(q)$ is given by $l$, that at $q$ is given by $r_q$, and the equality $l=r_qm_q$ has to hold.
\end{remark}

\end{definition}

>>>>

\subsection{Relative logarithmic maps to expanded pairs} We recall a number of definitions from \cite{ChenDegeneration}; see also \cite{KimLog}.

@@

Finally, fix an $n$-tuple of non-negative contact orders $(c_1,\ldots,c_n)$.
\begin{definition}
 A \emph{relative log prestable map} with tangency condition $\mathbf{c}$ is given by a diagram of log morphisms:
 \bcd
 ((C,\mathbf p),\pM_C)\ar[rr,"f"]\ar[dr] & & ((W,D),\pM_W)\ar[rr,"\pi_X"]\ar[dl] & & (X,Y)\times S \\
  & (S,\pM_S) & & &
 \ecd
 \begin{enumerate}
  \item $(C,\mathbf p),\pM_C)\to (S,\pM_S)$ is a minimal log twisted curve;
  \item $((W,\pM_W)\to (S,\pM_S),(W,D)\to(X,Y))$ is an extended twisted log expansion of $(X,Y)$;
  \item (\emph{corank condition}) \sidenote{split the definition into a more general one over LogSch and minimality?} for every geometric point $\bar{s}\in S$, the rank of \[\operatorname{Coker}(\pM^{W/S}_{S,\bar{s}}\to \pM_{S,\bar{s}})\] is equal to the number of non-distinguished nodes of $C_{\bar{s}}$; recall that a node of $C_{\bar{s}}$ is \emph{distinguished} if it is mapped under $f$ to the singular locus of $W_{\bar{s}}$.
  \item (\emph{log admissibility}) the morphism of log structures $f^\flat\colon f^*\pM_W\to\pM_C$ is simple at every distinguished node;
  \item (\emph{tangency condition}) the underlying map $\underline{f}\colon C\to (W,D)$ is non-degenerate, namely no component of $C$ is mapped entirely into $D$, and locally around every marking $p_i$ we can find a chart
  \bcd
  \NN\ar[r,"\cdot c_i"]\ar[d] & \NN\ar[d] \\
  f^*\pM^D\ar[r] & \pM^{p_i}
  \ecd
  for the restriction of $f^\flat$ to the $f^*\pM^D$ factor.
  \end{enumerate}
\end{definition}
\begin{definition}
 A relative log prestable map as above is \emph{stable} if for every geometric point $\bar{s}\in S$ the group of automorphisms $(\sigma,\tau)$ is finite, where:
 \begin{enumerate}
  \item $\sigma$ is an automorphism of $((C,\pM_C)\to (S,\pM_S))_{\bar{s}}$ preserving the markings $\mathbf p$;
  \item $\tau$ is an automorphism of $((W,\pM_W)\to (S,\pM_S))_{\bar{s}}$ preserving $D$ and $\pi_X\colon W_{\bar{s}}\to X$;
  \item $\tau\circ f_{\bar{s}}=f_{\bar{s}}\circ \sigma$.
 \end{enumerate}
\end{definition}
\begin{remark}
 The corank and simplicity condition imply the following: the rank of the locally free log structure $\pM_S$ at $\bar{s}$ amounts to the number of irreducible components of $W_{\bar{s}}^{\text{sing}}$ plus the number of non-distinguished nodes of $C_{\bar{s}}$; all the smoothing parameters of the distinguished nodes mapping to the same component of $W_{\bar{s}}^{\text{sing}}$ have been identified \emph{up to twisting}. In facts, at a distinguished node $q$ of $C$, letting $R$ be the local ring of $S$ at $\pi(q)$, we can find \'{e}tale local charts
 \bcd
 \NN^2\oplus_{\NN} \pM_{S}\ar[r,"{(id,\cdot m_q)}" below]\ar[d] & \NN^2\oplus_{\NN} \pM_{S}\ar[d] \\
 f^*\pM_W \ar[r,"f^\flat" below]\ar[d] & \pM_C\ar[d] \\
 R[z_1,\ldots,z_{\dim(X)-1},x,y]/(xy-t)\ar[r,"\left(\substack{x\mapsto a^{m_q}\\ y\mapsto b^{m_q}\\ t\mapsto s^{m_q}}\right)" below=.2cm] & R[a,b]/(ab-s)
 \ecd
 where the twist at $f(q)$ is given by $l$, the one at $q$ is given by $r_q$, and the equality $l=r_qm_q$ has to hold.
\end{remark}

\subsection{The desingularisation via central alignments} The following is a variation on \cite[\S 3.3]{RSPW2}.
